---
title: "Roofline"
author: "Fabian Wüthrich"
date: "5/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intel Core i7-8650U

- [Intel Spec](https://ark.intel.com/content/www/us/en/ark/products/124968/intel-core-i7-8650u-processor-8m-cache-up-to-4-20-ghz.html)
- [Wikichip CPU](https://en.wikichip.org/wiki/intel/core_i7/i7-8650u)
- [Wikichip Microarch](https://en.wikichip.org/wiki/intel/microarchitectures/kaby_lake)

## Peak Performance (pi)

Assume Port 0 and 1 can handle FP FMAs

see https://en.wikichip.org/wiki/intel/microarchitectures/skylake_(client)#Scheduler and Intel Opt. Chapter 11.15

- without SIMD:   2 FMAs/cycle => 4 flops/cycle
- with SIMD:      4 doubles * 4 flops/cycle => 16 flops/cycle

## Bandwidth (beta)

Roofline paper used special STREAM benchmark (see cacm_roofline_appendix_a.pdf)

Püschel used Intel MKL copy routine (see applying_roofline.pdf Section V. G)

_Which beta should we choose?_

Wikichip: 8 byte/cycle (see https://en.wikichip.org/wiki/intel/microarchitectures/kaby_lake#Memory_Hierarchy)

Intel Spec says 37.5 GB/s for two channels (18.75 GB/s for one core) => 9.8684 byte/cycle

Pueschel's law 60% of spec: (18.75 GB/s * 0.6) / 1.9 GHz ~ 6 bytes/cycle

STREAM (single core): 17.1655 GB/s => 9.0345 bytes/cycle

ERT (with SIMD): 17.13 GB/s => 9.0158 bytes/cycle

==> I think 9 bytes/cycle are ok on my machine :)

## Algorithm Model

- Work W(n) and Runtime T(n) empirical by instrumenting code
- Data movement Q(n) only asymptopically (load image once i.e. 8*n²)

## Plots

```{r plot, echo=FALSE}
library(tidyverse)

pi_scalar <- 4
pi_simd <- 16
beta <- 9

roofline <- function(x) ifelse(x*beta<pi_scalar, x*beta, pi_scalar)

benchmark <- tibble(
  opt =    c('baseline', 'baseline',  'baseline',  'baseline' ),
  n =      c(64,         256,         512,         1024       ),
  flops =  c(2.0804e+07, 1.07571e+09, 4.76046e+09, 3.9249e+10 ),
  cycles = c(4.6464e+07, 2.30593e+09, 9.59347e+09, 7.53262e+10),
  bytes = 8*n^2,
  perf = flops/cycles,
  op_int = flops/bytes,
)

ggplot(mapping = aes(x = op_int, y = perf, color = opt), data = benchmark) + 
  geom_point() +
  geom_line() +
  geom_hline(yintercept = pi_scalar, color = 'darkgrey') +
  geom_hline(yintercept = pi_simd, color = 'darkgrey') +
  geom_abline(slope = 1, intercept = 3.2, color = 'darkgrey') +
  # uncomment to see exact roofline function
  #stat_function(fun = roofline) +
  scale_y_continuous(trans='log2', limits = c(NA, 20)) +
  scale_x_continuous(trans='log2', limits = c(0.08, NA)) +
  geom_text(aes(x=1400, y=pi_scalar, label = "Peak scal. (4 Flops/Cycle)"), colour = 'darkgrey', vjust=-1) +
  geom_text(aes(x=1300, y=pi_simd, label = "Peak vec. (16 Flops/Cycle)"), colour = 'darkgrey', vjust=-1) +
  geom_text(aes(x=0.17, y=1.4, angle = 60, label = "Bandwidth (9 Bytes/Cycle)"), colour = 'darkgrey', vjust=-1) +
  labs(
    y = 'Performance [Flops/Cycle]',
    x = 'Operational Intensity [Flops/Byte]'
  ) +
  theme(legend.position = "none")
```

```{r save, echo=FALSE}
ggsave("roofline.pdf", width = 7.3, height = 4.5)
```

## Appendix

### STREAM one core

- Expected bandwidth ~18.75 GB/s
- Increasing STREAM_ARRAY_SIZE gives compiler error

```
$ gcc -Ofast  -DSTREAM_ARRAY_SIZE=100000000 stream.c -o stream
$ ./stream 
-------------------------------------------------------------
STREAM version $Revision: 5.10 $
-------------------------------------------------------------
This system uses 8 bytes per array element.
-------------------------------------------------------------
Array size = 100000000 (elements), Offset = 0 (elements)
Memory per array = 762.9 MiB (= 0.7 GiB).
Total memory required = 2288.8 MiB (= 2.2 GiB).
Each kernel will be executed 10 times.
 The *best* time for each kernel (excluding the first iteration)
 will be used to compute the reported bandwidth.
-------------------------------------------------------------
Your clock granularity/precision appears to be 1 microseconds.
Each test below will take on the order of 93592 microseconds.
   (= 93592 clock ticks)
Increase the size of the arrays if this shows that
you are not getting at least 20 clock ticks per test.
-------------------------------------------------------------
WARNING -- The above is only a rough guideline.
For best results, please be sure you know the
precision of your system timer.
-------------------------------------------------------------
Function    Best Rate MB/s  Avg time     Min time     Max time
Copy:           17165.5     0.093353     0.093210     0.093527
Scale:          12825.2     0.125195     0.124754     0.126527
Add:            14243.0     0.169058     0.168504     0.170513
Triad:          14066.8     0.170994     0.170614     0.171486
-------------------------------------------------------------
Solution Validates: avg error less than 1.000000e-13 on all three arrays
-------------------------------------------------------------
```

=> got 17.2 GB/s i.e. 17.2 GB/s / 1.9 GHz = 9.05 bytes per cycle (turbo boost turned off)

### STREAM all cores

```
$ gcc -Ofast -fopenmp -DSTREAM_ARRAY_SIZE=100000000 stream.c -o stream_omp
$ ./stream_omp 
-------------------------------------------------------------
STREAM version $Revision: 5.10 $
-------------------------------------------------------------
This system uses 8 bytes per array element.
-------------------------------------------------------------
Array size = 100000000 (elements), Offset = 0 (elements)
Memory per array = 762.9 MiB (= 0.7 GiB).
Total memory required = 2288.8 MiB (= 2.2 GiB).
Each kernel will be executed 10 times.
 The *best* time for each kernel (excluding the first iteration)
 will be used to compute the reported bandwidth.
-------------------------------------------------------------
Number of Threads requested = 8
Number of Threads counted = 8
-------------------------------------------------------------
Your clock granularity/precision appears to be 1 microseconds.
Each test below will take on the order of 80806 microseconds.
   (= 80806 clock ticks)
Increase the size of the arrays if this shows that
you are not getting at least 20 clock ticks per test.
-------------------------------------------------------------
WARNING -- The above is only a rough guideline.
For best results, please be sure you know the
precision of your system timer.
-------------------------------------------------------------
Function    Best Rate MB/s  Avg time     Min time     Max time
Copy:           20202.3     0.080604     0.079199     0.086724
Scale:          13990.7     0.115141     0.114362     0.116096
Add:            14049.4     0.172711     0.170826     0.175254
Triad:          14439.4     0.166823     0.166212     0.167873
-------------------------------------------------------------
Solution Validates: avg error less than 1.000000e-13 on all three arrays
-------------------------------------------------------------
```

=> got 20.2 GB/s but now we have multiple channels...

### ERT scalar

see doc/roofline/roofline-toolkit/Empirical_Roofline_Tool-1.1.0/Results/skylake-scalar/Run.001/roofline.json

```
"DRAM",         13.94
"FP64 GFLOPs",  5.45
```

### ERT SIMD

see doc/roofline/roofline-toolkit/Empirical_Roofline_Tool-1.1.0/Results/skylake-simd/Run.001/roofline.json

somehow I have caches here don't ask me why...

```
"L1",           37.43             
"L2",           29.95
"DRAM",         17.13
"FP64 GFLOPs",  9.37
```

### Benchmark Baseline

```
$ ./build/baseline -b -c -f images/lena_64.gray 
PARAMETERS
	File: images/lena_64.gray
	Error: 350
VERIFICATION phase
	PSNR (dB): 26.9019 (at least 25 dB / higher is better)
	✔ verification succeeded
WARMUP phase
	performing function for at least 1e+08 cycles
	2 runs needed
BENCHMARK phase
	cycles (median): 4.6464e+07
	flops: 2.0804e+07
	perf [flops/cycle(median)]: 0.447744

$ ./build/baseline -b -c -f images/lena_256.gray 
PARAMETERS
	File: images/lena_256.gray
	Error: 350
VERIFICATION phase
	PSNR (dB): 25.7653 (at least 25 dB / higher is better)
	✔ verification succeeded
WARMUP phase
	performing function for at least 1e+08 cycles
	2 runs needed
BENCHMARK phase
	cycles (median): 2.30593e+09
	flops: 1.07571e+09
	perf [flops/cycle(median)]: 0.4665

$ ./build/baseline -b -c -f images/lena_512.gray 
PARAMETERS
	File: images/lena_512.gray
	Error: 350
VERIFICATION phase
	PSNR (dB): 25.7578 (at least 25 dB / higher is better)
	✔ verification succeeded
WARMUP phase
	performing function for at least 1e+08 cycles
	2 runs needed
BENCHMARK phase
	cycles (median): 9.59347e+09
	flops: 4.76046e+09
	perf [flops/cycle(median)]: 0.496219

$ ./build/baseline -b -c -f images/grey-parrot_1024.gray 
PARAMETERS
	File: images/grey-parrot_1024.gray
	Error: 350
VERIFICATION phase
	PSNR (dB): 25.4697 (at least 25 dB / higher is better)
	✔ verification succeeded
WARMUP phase
	performing function for at least 1e+08 cycles
	2 runs needed
BENCHMARK phase
	cycles (median): 7.53262e+10
	flops: 3.9249e+10
	perf [flops/cycle(median)]: 0.521053
```