\section{Background}\label{sec:background}
To introduce the algorithm, we first touch a glimpse of the underlying theory, then show how one can apply
it to compress images and then finally show practical considerations for the implementation.
For this chapter we use the notation from Fisher \cite{fisher2012}.

\mypar{Iterated function systems (IFS)}
Fractal image compression builds on the theory of iterated function systems. For completeness,
we provide a short introduction to this rich topic.
\begin{definition}[Contractivity of functions]
    A function $f$ on a metric space $X$ with metric $d$ is contractive, when there exists some $0 \leq k < 1$
    such that for all $x,y \in X: d(f(x), f(y)) \leq k \cdot d(x,y)$.
\end{definition}
As an informal example, when $f$ is a contractive function on $\R$, then mapping two numbers brings them \textit{closer} together,
i.e. their distance gets smaller.
\begin{definition}[Iterated function system]
    An iterated function system (IFS) is a set of $n$ functions $$\{f_i:X \to X | i=1,...,n\}$$
    where each $f_i$ is contractive and $X$ is a metric space with the metric $d$.
\end{definition}
\begin{definition}[Hutchinson operator]
    The Hutchinson operator for an IFS is the function
    \begin{align*}
        F \colon 2^X &\to 2^X\\
        x &\mapsto \bigcup_{i=1}^n f_i(x)
    \end{align*}
\end{definition}
With $F^{\circ n}(x)$, we denote the iterative application of $F$ $n$ times on its input $x$, e.g. $F^{\circ 2}(x) = F(F(x))$.
\begin{definition}[Attractor of an IFS]
    Let $A \in X$. When $F(A)=A$, then $A$ is called attractor.
\end{definition}
The following theorem about attractors has been proved by Hutchinson in \cite{hutchinson1981fractals} and is
the essential theoretical part how fractal image compression works.
\begin{theorem}[Contractive mapping fixed-point theorem]\label{theorem-fixpoint}
    For an IFS on a set $X$ it holds that:
    \begin{enumerate}
        \item There always exists a unique attractor $A \in X$
        \item For any nonempty set $S_0 \subset X$ it holds that $A = \lim_{n \to \infty} F^{\circ n}(S_0)$
    \end{enumerate}
\end{theorem}
Note that in the theorem above, we require that $X$ is a compact set, which holds for fractal image compression.

\mypar{From IFS to image compression}
The key idea in fractal image compression is that one computes a set of contractive functions $w_i \in \boldsymbol{W}$ (in this context called transformations) on the image.
From theorem \ref{theorem-fixpoint}, we then know that some unique attractor $A$ exists. If the transformations $\boldsymbol{W}$ are chosen in a way such that the attractor $A$ is the original image,
one only needs to store the transformations. For decompression, one can iteratively apply all transformations on any initial starting image, which then converges to the original image to be compressed.\\
A transformation $w_i$ is defined on a contiguous source region (called domain block) and a contiguous target region (range block) of the image.
The transformation maps the pixels of the domain block to the pixels of the range block and then applies a brightness and saturation adjustment.\\
A high-level algorithmic approach for compression is then:
\begin{enumerate}
    \item Partition the image into range blocks $\boldsymbol{R}$ and domain blocks $\boldsymbol{D}$.
    \item For each range block $R_i \in \boldsymbol{R}$, find a transformation $w_i$ with domain block $D_i \in \boldsymbol{D}$
            such that $w_i(D_i)$ approximates the pixel values of block $R_i$ best.
\end{enumerate}
\mypar{Computing transformations}
A grayscale image can be interpreted as a function
\begin{align*}
    f \colon I^2 &\to I\\
    (x,y) &\mapsto z = f(x,y)
\end{align*}
where $I$ is the interval $[0,1]$. Specifically, an image is a set of three-tuples $(x,y,z) \in I^3$, where $x,y$ are positions and $z$ is the grayscale value.
The image of a range (or domain) block $R_i$ is then defined by $f \cap (R_i \times I)$.
A transformation $w_i$ from the image of a domain block $D_i$ to the image of a range block $R_i$ can then be described as a linear affine transformation
$$
w_i(D_i) = \bigcup_{(x,y,z) \in f \cap (D_i \times I)} \begin{pmatrix} a_i & b_i & 0 \\ c_i & d_i & 0 \\ 0 & 0 & s_i \end{pmatrix} \begin{pmatrix} x\\y\\z \end{pmatrix} + \begin{pmatrix} e_i\\f_i\\o_i \end{pmatrix}
$$
where $a_i, b_i, c_i, d_i, e_i, f_i$ map \textit{location} (which can involve rotations) and $s_i, o_i$ correspond to contrast and brightness adjustments.
To ensure that $w_i$ is contractive, $D_i$ has to be larger than $R_i$ (in terms of region sizes) and $0 \leq s_i < 1$.

Now we define a metric to compare two images (or two blocks). Usually, the root mean square error (RMS) metric is used:
$$
d_{RMS}(f,g) = \sqrt{\int_{I^2} (f(x,y)-g(x,y))dxdy}
$$
For a range block $R_i$ and domain block $D_i$ we now seek the transformation $w_i$, such that $d_{RMS}(w_i(D_i), f \cap (R_i \times I))$ is minimal, i.e.
\begin{equation} \label{eq:min-problem}
    w_i = \argmin_{w} \left( d_{RMS}(w(D_i), f \cap (R_i \times I)) \right)
\end{equation}
Calculating $w_i$ is a minimization problem, because we may choose $s_i$ and $o_i$ in order to minimize the error.\\
The set of all transformations is a partitioned iterated function system (PIFS). The only difference to an IFS is that the transformations are restricted
to \textit{blocks} (partitions) of the image. However, the underlying theory still holds \cite{fisher2012}.

\mypar{Practical implementation} \label{par-practical-implementation}
In practice an image consists of pixels, where each pixel has a grayscale value between 0 and 1 and integer coordinates.
\notefabian{Ist unser Bild nicht 0...255?}
The minimization problem in equation \ref{eq:min-problem} can be solved using least squares regression, as done in \cite{github-python}.
Fisher uses an analytical solution in \cite{fisher2012} to compute the adjustments directly. Let $b_1, ..., b_n$ be the $n$ pixel values of the range block $R_i$
and $a_i,...,a_n$ be the $n$ pixel values of the downsampled and rotated domain block. We then seek $s$ (saturation) and $o$ (brightness) such that
$$
R = \sum_{i=1}^n (s \cdot a_i + o - b_i)^2
$$
is minimal. An analytical solution for $s,o,R$ is:
\begin{align}
    s &= \frac{n \sum_{i=1}^n a_i b_i - \sum_{i=1}^n a_i \sum_{i=1}^n b_i}{n \sum_{i=1}^n a_i^2 - \left(\sum_{i=1}^n a_i \right)^2} \label{contrast}\\
    o &= \frac{1}{n} \left( \sum_{i=1}^n b_i - s \sum_{i=1}^n a_i \right) \label{brightness}\\
    R &= \frac{1}{n} \left[ \sum_{i=1}^n b_i^2 + s \left( s \sum_{i=1}^n a_i^2 - 2 \sum_{i=1}^n a_i b_i + 2o \sum_{i=1}^n a_i \right) \right. \notag\\
    &\text{\hspace{1cm}} \left. + o \left(no - 2 \sum_{i=1}^n b_i \right)  \right]\label{error}
    \end{align}
Thus, to compute an optimal transformation $w_i$ and the error for a given range block $R_i$ and given domain block $D_i$, one needs to compute the sums above
as efficient as possible.

\mypar{Image partitioning}
There are many ways to partition an image into range and domain blocks. We require that the range blocks cover the full image and do not overlap, otherwise the
decompressed image would have uncovered regions. The domain blocks represent the search space for self-similarity and may overlap.
Increasing the domain block pool $\boldsymbol{D}$ may lead to better transformations but can increase the compression performance significantly.

A simple approach is exhaustive block mapping, where each range block is compared with each domain block. Blocks can be of various sizes and shapes 
(quadratic, rectangular, etc.), but domain blocks must be larger than range blocks to satisfy the contractivity requirement. For example the code in \cite{github-python}, 
uses quadratic range blocks of size $s \times s$ and domain blocks of size $k\cdot s \times k \cdot s$ for some integer constant $k > 1$.

In practice, quadtree partitioning is considered a reasonable reference point for more advanced partitioning schemes \cite{fisher2012}.
It dynamically adapts range and domain block sizes by a predefined error threshold $\epsilon$.
We start with some initial (quadratic) range blocks of size $s \times s$, and domain blocks of size $2s \times 2s$.
When there is a range block for which all domain blocks exceed the threshold $\epsilon$, we partition
the range block into 4 smaller blocks of size $s/2 \times s/2$ and try to cover them again with domain blocks of size $s \times s$.
This scheme has the effect that (potentially larger) homogeneous regions of the image are covered with few transformations while image
regions with details are covered with transformations which map smaller blocks.
Usually, one defines also a maximum depth of the quadtree as described in \cite{fisher2012}.

\mypar{Cost analysis} 

We provide a pessimistic upper bound on the cost for our
algorithm where we assume that no transformation will yield an error below
the threshold and thus the only termination criteria is the maximum quadtree
depth. In practice, the cost will be (much) lower assuming a reasonable error
threshold $\epsilon$. As a cost metric, the number of floating point operations
are counted.

\begin{equation} \label{eq:cost}
C = N_{mult} + N_{div} + N_{add} + N_{sub}
\end{equation}

Let $m$ be the maximum quadtree depth and let $s$ be the width of the square
input image. We assume that we start quadtree with range blocks of size
$\frac{s}{2}\times\frac{s}{2}$ and one domain block of size $s \times s$. For
quadtree depth $i \in \{1, \dots, m\}$, let $R^{(i)}$ be the set of range
blocks, $D^{(i)}$ the set of domain blocks, $S_{R}^{(i)}$ the number of pixels
of one range block and $S_{D}^{(i)}$ the number of pixels of one domain block.
For a given domain block and range block in quadtree depth $i$, let $\alpha_{i}$
denote the cost to compute four transformations $w_i^{0^}$, $w_i^{90^}$, 
$w_i^{180^}$, $w_i^{270^}$  and their errors. Each transformation $w_i^{\gamma^}$ 
represents a rotation of the domain block by $\gamma$ degrees.

To calculate $\alpha_{i}$, we need to compute all sums for brightness,
contrast and error (equations \eqref{contrast}, \eqref{brightness} and
\eqref{error}) and apply the formulas. In addition, we need to downscale and
rotate the domain block.

Note that the sums $\sum_{i=1}^{n} a_i$,$\sum_{i=1}^{n} a_i^2$,
$\sum_{i=1}^{n} b_i$,$\sum_{i=1}^{n} b_i^2$ are the same for all four
rotations and $\sum_{i=1}^{n} a_i b_i$ has to be computed for each rotation.

Downsampling can be done in $S_{D}^{(i)}$ flops because we have to aggregate the
average of four pixels of the domain block into one pixel which requires 4 flops
for every square of four pixels.

\begin{align*}
  \alpha_i &= S_{D}^{(i)} + 4 \cdot (2\cdot S_{R}^{(i)}-1) + (6\cdot S_{R}^{(i)}-4) +  26\\
           &= S_{D}^{(i)} + 14\cdot S_{R}^{(i)} + 18 \\
           &= 18\cdot S_{R}^{(i)} + 18 \\
         % &= 4 \cdot \frac{S^2}{2^{2i}} + 14 \cdot \frac{S^2}{2^{2i}} - 8 \\
           &= 18 \cdot \frac{s^2}{2^{2i}} + 18
\end{align*}

Thus, we get the following upper bound on the cost:

\begin{align*}
C &= \sum_{i=1}^m |R^{(i)}| \cdot |D^{(i)}| \cdot \alpha_i \\
  &= \sum_{i=1}^m 2^{2i} \cdot 2^{2i-2} \cdot (18 \cdot \frac{s^2}{2^{2i}} + 18) \\
  &= \frac{6}{5}\cdot (4^{m}-1)\cdot (4^{m+1} + 5\cdot s^{2} + 4)
\end{align*}

We see that to cost of the algorithm is exponential in the quadtree depth $m$
and polynomial in image width $s$.

% TODO: cost -> running time / time complexity?
% i.e. it's time complexity is $\mathcal{O}(S^24^{M} - 4^{2M+1})$.

% TODO: but is in chapter 3 or even 4
% Our experiments showed that this upper bound is way
% too pessimistic and the exponentiality in $M$ is practically not a problem
% given a reasonable $S,M,\epsilon$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Give a short, self-contained summary of necessary
% background information on the algorithm or application that you then later optimize including a cost analysis.

% For example, assume you present an
% implementation of FFT algorithms. You could organize into DFT
% definition, FFTs considered, and cost analysis. The goal of the
% background section is to make the paper self-contained for an audience
% as large as possible. As in every section
% you start with a very brief overview of the section. Here it could be as follows: In this section
% we formally define the discrete Fourier transform, introduce the algorithms we use
% and perform a cost analysis.

% \mypar{Discrete Fourier Transform}
% Precisely define the transform so I understand it even if I have never
% seen it before.

% \mypar{Fast Fourier Transforms}
% Explain the algorithm you use.

% \mypar{Cost Analysis}
% First define you cost measure (what you count) and then compute or determine on other ways the
% cost as explained in class. In the end you will likely consolidate it into one number (e.g., adds/mults/comparisons) but be aware of major imbalances as they affect the peak performance..

% Also state what is known about the complexity (asymptotic usually)
% about your problem (including citations).
