\section{Contributions of Team Members}

\hspace{\parindent}\mypar{Jonas} Focused on the baseline implementation and the following scalar
optimizations (ILP, removal of structs, removal of pointer chasing and
decreasing of memory reallocation). Helped in prototyping SIMD (especially the
four-domain-blocks-a-time approach)
and contributed to the solution which increased SIMD performance regarding 90 and 270 degree rotations.\\
Wrote a tool to simulate cache access in order to determine tighter memory bounds in the roofline plot.\\
Experimented with block-wise domain/range block traversal.

\mypar{Pascal} Added argument processing, file I/O and some parts in the
brightness and contrast computation of the baseline implementation. Added test
data and prepared performance and runtime analysis. Did bottleneck analysis and
contributed to the scalar optimizations (e.g. merging loops, more efficient
precomputations). Contributed to the SIMD implementation (e.g. eliminate
expensive instructions, scaling blocks, version without rotation). Tested
compiler flags including ICC.

\mypar{Fabian} Added rotations to baseline. Created roofline plot and compared baseline to reference
implementation. ILP and index optimizations for $\sum_{i=1}^n a_i b_i$ calculation.

Implemented parts of four-domain-blocks-a-time SIMD (tricky part was error calculation with
\verb|blendv| and mask) and participated in bringing the two SIMD approaches together.
Helped increasing performance regarding 90 and 270 degree rotations.

\mypar{Janis} Changed baseline from C\texttt{++} to C and started with the implementation of a fast queue.
Due to personal problems he withdrew from the project.
