% IEEE standard conference template; to be used with:
%   spconf.sty  - LaTeX style file, and
%   IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------

\documentclass[letterpaper]{article}

% custom setup
\input{setup.tex}

% set to true or false to enable or disable notes
\newboolean{shownotes}
\setboolean{shownotes}{true}

% Title.
% ------
\title{Performance Optimized Fractal Image Compression using Quadtree}
%
% Single address.
% ---------------
\name{Jonas Hansen, Pascal Huber, Fabian Wüthrich\thanks{Thanks to Janis Peyer for his contributions}}
\address{Department of Computer Science\\ ETH Zurich, Switzerland}

% For example:
% ------------
%\address{School\\
%		 Department\\
%		 Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%		 {School A-B\\
%		 Department A-B\\
%		 Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%		 while at ...}}
%		 {School C-D\\
%		 Department C-D\\
%		 Address C-D}
%

\begin{document}
%\ninept
%
\maketitle
%
%

% space efficient align
\setlength{\abovedisplayskip}{1ex}
\setlength{\belowdisplayskip}{1ex}
% \setlength{\abovedisplayshortskip}{0pt}
% \setlength{\belowdisplayshortskip}{0pt}

\begin{abstract}
  Fractal image compression is a lossy image compression method, which yields
  good compression ratios and quality at the cost of high encoding times. In
  this work, we started with our own implementation of a fractal compression
  scheme using the popular quadtree partitioning and exhaustive search for
  self-similarity in the image.

  We analyzed our baseline implementation with the roofline model and used
  profiling to identify various performance bottlenecks. With these insights we
  implemented precomputations and removed complex data structures to improve
  performance. In addition, we optimized the code for instruction level
  parallelism and used FMA instructions for an optimal pipeline utilization.
  Then we vectorized the code using AVX2 intrinsics and changed the memory
  layout to avoid expensive gathering instructions.

  Our optimized version without vectorization led to a 4x speedup whereas we
  observed a speedup of up to 8x with SIMD. Additionally, we improved the
  runtime of the encoding part significantly. For an image of the size
  $2048 \times 2048$ pixels the compression time was reduced from 2 hours to 2.5
  minutes.

  \notepascal{den letzten satz hätte ich etwas genereller formuliert - der
    klingt ein wenig zu gut...}
\end{abstract}

\input{01_introduction.tex}
\input{02_background.tex}
\input{03_your_method.tex}
\input{04_results.tex}
\input{05_conclusions.tex}
\input{06_contributions.tex}

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: bibl_conf). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{bibl_conf}

\end{document}
